---
title: "Metapopulation Migration/Extinction Model"
author: Emma Ochsner and Lena Witschi
output: html_notebook
---

# Introduction

This is the code for our Model for Metapopulations in SLiM. We first softcode migration rate in R and how many reps we choose and which rates.

## Generating Data in SLiM with Parameters from R

```{r}

rm(list=ls(all=TRUE)) ## clear R environment

options(scipen = 999) ## scientific notation penalty

setwd("~/Documents/GitHub/THEE_Practical_Project/Russ")

## choose number of reps for each parameter set
reps <- 5


## Choose the different parameter values we are interested in (migration rate)
all_m <- c(0.1, 0.2, 0.3, 0.4)
repetition <- c(1, 2, 3, 4, 5)

#make a for loop for all the different values we want to test for m 
for(m in all_m){
  
    for(r in 1:reps){
  slim_code_block <- paste0("slim -d migration=",m," -d repetition=", r," metapop_script_basic_test.slim")
  #slim_code_block -> might not be needed...to still be tested
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}
  }

#outputs a csv file with all our data :) - (1) the mutation’s id property, (2) the identifier of its mutation type, (3) its position, (4) its selection coefficient, (5) its dominance coefficient, (6) origin subpopulation identifier, (7) origin tick, and (8) prevalence.


```

## Reading in and analysing Data

In this R block, we start with reading in the Data we generated before. We then made a data frame to store all the metrics we need to start further analysis. This Data frame is saved as Example_result. 
We calculate the Heterozygosities within the populations for all ten patches and then fill in the Dataframe. 

```{r}
## all SLiM output files
slim_out_files1 <- list.files(pattern = "slim_muts_output")
slim_out_files <- slim_out_files1[!grepl("extinction", slim_out_files1)]

#install.packages("hierfstat")
library("hierfstat")

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run
results_df <- data.frame(array(NA, dim = c(length(slim_out_files),13), dimnames = list(c(),c("m","reps","seed","p1", "p2", "p3", "p4", "p5", "p6", "p7", "p8", "p9", "p10"))))

## loop over all SLiM results files
for(i in 1:length(slim_out_files)){ 
  
  single_file <- slim_out_files[i] ## this is the name of a single SLiM result file
  
  get_the_m <- as.numeric(gsub(".*_migration =|_seed.*", "", single_file))## here we use gsub() to get the value of 'm' that we saved in the name of the results file
   get_the_reps <- as.numeric(gsub(".*_reps_|_migration.*", "", single_file)) ## do the same for reps
  
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) ## and for the seed....
  
  single_file_df <- read.csv(slim_out_files[i], sep=" ", header=F, stringsAsFactors=F) ## read the data in the SLiM results file into R
  
  ## adding in column names
  colnames(single_file_df) <- c("","Max Number of Gen","Number of Generations", "", "Deme", "mut id", "mut typ", "position on gemome", "fitness", "dominance", "original deme", "gen mut occured", "number of indiv. with mut")

  ##Create a new column with allele frequencies (Add allele freq. in a separate row (for each file))
  ##1000 is K -> So for allele freq. you have to calculate x/2000 -> consider we have diploid individuals and K=1000 -> 2000 alleles
   single_file_df$Allelefreq <- single_file_df[,13]/2000
   
  ##calculate heterozygosity per locus (2pq -> p is allele freq and q = 1-p) (within deme)
  single_file_df$Heterozygosity_l <- 2*single_file_df[,14]*(1-single_file_df[,14])
  
  ###Next step: convert to genome wide heterozygosity (within deme)
  #na.rm = TRUE ensures that any missing values (NA) in the heterozygosity column are ignored during the calculation.
  
  #genome wide heterozygosity for the sub populations
  for (x in 1:10){
  sel_row_p <- single_file_df[,5] == paste0("p",x)
  vektor_p_pol <- single_file_df[sel_row_p,15]
  vektor_p_0 <-rep(0,10000 - length(vektor_p_pol))
  h_p <- mean(c(vektor_p_pol,vektor_p_0), na.rm = TRUE)
  results_df[i,paste0("p",x)] <- h_p
  }
  

  ##Fill in the Data Frame  
  
  results_df[i,"m"] <- get_the_m ## record the value of 'u' for this particular results file
  
  results_df[i,"reps"] <- get_the_reps ## record the value of 'r' for this particular results file
  
  results_df[i,"seed"] <- get_the_seed ## record the seed for this particular results file (this lets us keep track of individual reps of a give parameter combination), ie, maybe we did 5 reps of u=1e-8 and r=1e-8, so we will have 5 rows for these specific parameter values
  
}


## choose working directory
## write your results to disc
setwd("~/Documents/GitHub/THEE_Practical_Project/Russ")
write.csv(results_df, "Example_Results.csv", row.names = F)


```


## First Analysis 

Here we first read in our result file before starting to calculate statistical metrics....

````{r}

##Read in the csv file
Example_results <- read.csv("Example_Results.csv")

##calculate the Means over the 10 subpopulations
row_mean <- rowMeans(Example_results[,4:13])
Example_results$mean <- row_mean

##calculate the Variance (might need to explain how in more detail)
sel_col <- Example_results[,4:13]
row_variances <- apply(sel_col, 1, var)
Example_results$variance <- row_variances

## Generating the plot (more info...)
y.y <- seq(min(Example_results$mean)*0.9, max(Example_results$mean)*1.1, length.out = 2)

par(mar=c(6,6,4,2))
plot( 0, type ="n", las = 1, ylab = "", xlab = "migration rates", cex.lab = 1.2, cex.main = 1.6,
     main = "Heterozygosity for different migration rates", xlim = c(0,1), ylim = y.y, lwd = 2)
title(ylab = "mean H", line = 4, cex.lab = 1.2)


##Calculate the Mean of the means of each rep.
mean_of_means <- numeric(length(all_m))
result_array <- array(NA, dim = c(length(all_m), 2), dimnames = list(c(), c("mean", "mig")))

for (i in seq_along(all_m)){
  m <- all_m[i]  # Get the current migration rate
  mig <- Example_results[, 1] == m  # Get the logical vector for the current migration rate
  mig_val <- as.numeric(Example_results[mig, 14])  # Extract the corresponding values from the 14th column
    mean_of_means[i] <- mean(mig_val, na.rm = TRUE)  # Compute mean, removing NA values
  # Store the mean and migration rate in the result array
  result_array[i, 1] <- mean_of_means[i]  # Fill the mean column
  result_array[i, 2] <- m  # Fill the migration rate column
}
result_array


#does not work yet.....
#mean_of_variance <- numeric(length(all_m))
#result_array_var <- array(NA, dim = c(3, 2), dimnames = list(c(), c("variance", "mig")))

#for (i in seq_along(all_m)){
  #m <- all_m[i]  # Get the current migration rate
  #mig_var <- Example_results[, 1] == m  # Get the logical vector for the current migration rate
  #mig_val_var <- as.numeric(Example_results[mig_var, 15])  # Extract the corresponding values from the 14th column
   # mean_of_variance[i] <- mean(mig_val_var, na.rm = TRUE)  # Compute mean, removing NA values
  # Store the mean and migration rate in the result array
  #result_array_var[i, 1] <- mean_of_variance[i]  # Fill the mean column
  #result_array_var[i, 2] <- m  # Fill the migration rate column
#}
#result_array_var
#until here 

## adding our data into the generated plot
points(Example_results$m,Example_results$mean, lwd = 3, col = "pink")
lines(result_array[,2],result_array[,1],  lwd = 3, col = "red")

legend(x = 0.6, y = 0.0013, legend = c("Mean H for Subpopulations", "Mean of all Means"), lwd = c(3,3), cex = 0.75, col = c("pink", "red"), bty = "n")


````

# Plot1: Mean H for subpop for different migration rates
Since the mean of the variances is so small, we do not show it in this plot. 